---
title: "homework-4"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{homework-4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, results='hide'}
library(BIS557)
library(reticulate)
use_condaenv("r-reticulate")
library(casl)
```


Part.1

ridge_py() implements a numerically-stable ridge regression that takes into account colinear (or nearly
colinear) regression variables, using Python code. The coefficients match the ones from the R implementation.

```{r}
library(BIS557)
#python implementation
fit_py <- ridge_py(Sepal.Length ~ .- Species, iris, contrasts = list(Species = "contr.sum"), lambda = 0.5)
fit_py$coefficients
#r implementation
fit_r <- ridge(Sepal.Length ~ .- Species, iris,
               contrasts = list(Species = "contr.sum"), lambda = 0.5)
fit_r$coefficients
```

Part.2

A function named lm_ofc() was created and it fits a linear model by reading in contiguous rows from a data frame. QR decomposition is used in this implementation. The results are close to the ones of the lm() function.

```{r}
set.seed(1234)
library(BIS557)
n <- 1e6
d <- 5
df <- as.data.frame(matrix(runif(n * d, min = -1, max = 1), nrow = n, ncol = d))
names(df) <- c("y", "V1", "V2", "V3", "V4")
N <-200
beta <- matrix(0, nrow = N, ncol = d)
#select contiguous rows from a data frame
for (i in 1:N){
start <- (i-1) * n/N + 1
end <- i * n/N
df_new <- df[start:end, ]
fit_linear_model <- lm_ofc(y ~ ., df_new)
beta[i, ] <- fit_linear_model$coefficients
}
#fitted model
coefficients <- apply(beta, 2, mean)
coefficients
#lm results
fit_lm <- lm(y  ~ ., df)
fit_lm$coefficients
```

Part. 3

install_github("statsmaths/casl")

lasso() function fits a LASSO regression model in Python based on the soft-threshold function. The coefficients are the same as the "casl" results.

```{r}
library(BIS557)
data(iris)
form <- Sepal.Length ~ .; df <- iris
df_no_na <- model.frame(form, df)
X0 <- qr.Q(qr(model.matrix(form, df_no_na, contrasts.arg = list(Species = "contr.sum"))))
yname <- as.character(form)[2]
y0 <- matrix(df_no_na[,yname],ncol = 1)
#lasso results using casl package
fit_casl <- casl_lenet(X0, y0, lambda = 0.01, maxit = 1e5)
fit_casl
#python implementation
fit_py <- lasso(Sepal.Length  ~ ., iris, lambda = 0.01, contrasts = list(Species = "contr.sum"))
fit_py$coefficients
```

Part.4

For the final project I plan to implement a convolutional neural network that will be able to detect and classify dog breeds based on a training dataset. Backward propagation will be used and out-of-sample prediction accuracy will be calculated to estimate the effectiveness of this network.


Reference:

[1] https://numpy.org/doc/stable/reference/generated/numpy.linalg.svd.html

[2] https://numpy.org/doc/stable/reference/generated/numpy.sign.html

[3] Textbook P182 - P193
